---
sort: 3
title: "Alpha and Beta Diversity Analysis"
author: "XiaZhao"
date: "2021-04-13"
output: 
  md_document: 
    df_print: kable
    number_sections: yes
    toc: yes
---

[`Return`](./)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Requires

<details>
<summary>
<font size=4>Requires</font>
</summary>

```{r tidy=TRUE,results='hold',message=FALSE}

library(data.table)
library(tidyverse)
library(ggthemes)
library(ggsci)
library(ggpubr)
library(survminer)
library(survival)
library(survivalROC)
library(reshape2)
library(ggExtra)
library(cowplot)
library(GUniFrac)
library(ade4)
library(phangorn)
library(fpc)
library(vegan)
source("../R_function/colors.R")
"%ni%"<-Negate("%in%")
theme_set(theme_few())
options(stringsAsFactors = F)
```
</details>


# Alpha and Beta Diversity Analysis


## Description

**Ecosystem**

a community plus the physical environment that it occupies at a given time

**Alpha diversity**

the diversity within a particular area or ecosystem; usually expressed by the number of species (i.e., species richness) in that ecosystem

**Beta diversity**

a comparison of of diversity between ecosystems, usually measured as the amount of species change between the ecosystems

**Gamma diversity**

a measure of the overall diversity within a large region. Geographic-scale species diversity according to Hunter (2002: 448)


ref: Alpha, Beta, and Gamma Diversity. (2021, January 4). Retrieved April 19, 2021, from https://bio.libretexts.org/@go/page/17392

## Alpha diversity

###  Relationship Alpha diversity  and Response

```{r warning=F,results= 'hold'}
meta<-fread("../Data/Data/meta.csv",data.table = F)
alpha<-fread("../Data/Data/alpha_diversity.csv",data.table = F)[,c(1:4)]
data<-merge(select(meta,c(Samples,Response,Cycle,Site)),alpha,by="Samples")
data<-melt(data,id.vars = c("Samples","Response","Cycle","Site"),variable.name = "Index",value.name = "Value")
data_BL<-subset(data,Cycle=="BL"&Response!="NE")
data_treat<-subset(data,Cycle!="BL"&Response!="NE")

data_BL$Group<-paste0(data_BL$Site,"_",data_BL$Response)
data_BL$Index<-factor(data_BL$Index,levels = c("shannon","simpson","sobs"))

data_treat$Group<-paste0(data_treat$Site,"_",data_treat$Response)
data_treat$Index<-factor(data_treat$Index,levels = c("shannon","simpson","sobs"))

p1<-ggplot(subset(data_BL,Site=="Saliva"),aes(Response,Value))+
  geom_dotplot(aes(fill=Response,color=Response),binaxis = "y",stackdir = "center",dotsize = .8)+
  stat_compare_means(ref.group = "NR",label = "p.signif")+
  stat_summary(fun.data = mean_sdl,fun.args = list(mult=1),geom = "errorbar",width=.1,size=0.5)+
  stat_summary(fun = median,geom = "crossbar",color="black",size=.4,width=.4)+
  facet_wrap(~Index,scales = "free")+ theme_few()+
  scale_color_manual(values = c("#FF0000FF","#00FFFFFF"))+
  scale_fill_manual(values = c("#FF0000FF","#00FFFFFF"))+
  ggtitle("Saliva (Baseline)")+
  theme_few(base_size = 8)+
  theme(axis.title.x = element_blank())
p2<-ggplot(subset(data_BL,Site=="Stool"),aes(Response,Value))+
  geom_dotplot(aes(fill=Response,color=Response),binaxis = "y",stackdir = "center",dotsize = .8)+
  stat_compare_means(ref.group = "NR",label = "p.signif")+
  stat_summary(fun.data = mean_sdl,fun.args = list(mult=1),geom = "errorbar",width=.1,size=0.5)+
  stat_summary(fun = median,geom = "crossbar",color="black",size=.4,width=.4)+
  facet_wrap(~Index,scales = "free")+ theme_few()+
  scale_color_manual(values = c("#80FF00FF","#8000FFFF"))+
  scale_fill_manual(values = c("#80FF00FF","#8000FFFF"))+
  ggtitle("Stool (Baseline)")+
  theme_few(base_size = 8)+
  theme(axis.title.x = element_blank())
p3<-ggplot(subset(data_treat,Site=="Saliva"),aes(Response,Value))+
  geom_dotplot(aes(fill=Response,color=Response),binaxis = "y",stackdir = "center",dotsize = .8)+
  stat_compare_means(ref.group = "NR",label = "p.signif")+
  stat_summary(fun.data = mean_sdl,fun.args = list(mult=1),geom = "errorbar",width=.1,size=0.5)+
  stat_summary(fun = median,geom = "crossbar",color="black",size=.4,width=.4)+
  facet_wrap(~Index,scales = "free")+ theme_few()+
  scale_color_jama()+
  scale_fill_jama()+
  ggtitle("Saliva (Treat)")+
  theme_few(base_size = 8)+
  theme(axis.title.x = element_blank())

##use the treated stool samples without duplication
BL_treat<-fread("../Data/Data/16pt_BL_treat_pairs_stool.csv",data.table = F)
BL_treat_alpha<-merge(select(BL_treat,c(Samples,Group,Response)),alpha,by="Samples")
stool_treat<-subset(BL_treat_alpha,Group=="Treat")
stool_treat1<-melt(stool_treat[,-2],id.vars = c("Samples","Response"),value.name = "Value",variable.name = "Index")
stool_treat1$Index<-factor(stool_treat1$Index,levels = c("shannon","simpson","sobs"))
p4<-ggplot(stool_treat1,aes(Response,Value))+
  geom_dotplot(aes(fill=Response,color=Response),binaxis = "y",stackdir = "center",dotsize = .8)+
  stat_summary(fun.data = mean_sdl,fun.args = list(mult=1),geom = "errorbar",width=.1,size=0.5)+
  stat_summary(fun = median,geom = "crossbar",color="black",size=.4,width=.4)+
  stat_compare_means(ref.group = "NR",label = "p.signif",method = "wilcox.test")+
  facet_wrap(~Index,scales = "free")+
  scale_color_manual(values = c("#FF0099FF","#FF9900FF"))+
  scale_fill_manual(values = c("#FF0099FF","#FF9900FF"))+
  ggtitle("Stool (Treat)")+
  theme_few(base_size = 8)+ 
  theme(axis.title.x = element_blank())
```

```{r warning=F,fig.height=5,fig.width=7,dpi=600,fig.align='center',warning=F,out.width ="80%",results='hold'}
plot_grid(p1,p2,p3,p4, labels = c("A","B","C","D"), ncol =2, nrow = 2)
```

### Patients with both saliva and stool samples at baseline

```{r warning=F,fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
paired_meta<-fread("../Data/Data/stool_saliva_BL_paried.csv",data.table = F)
alpha_paired<-subset(alpha,Samples%in%paired_meta$Samples)
alpha_paired<-merge(paired_meta,alpha_paired,by="Samples")%>%filter(.,Response!="NE")
ggplot(alpha_paired,aes(Site,shannon,fill=Site))+
  geom_boxplot()+
  geom_line(aes(group=patientID))+
  geom_point(size=1)+
  theme_few(base_size = 8)+
  stat_compare_means(label = "p.signif")+
  facet_wrap(~Effect,scales = "free_x")+
  scale_fill_d3()
```

### Paired baseline saliva and stool samples

```{r warning=F,results= 'hold'}
load("../Data/Data/regaMicrcobiome.RData")
phylum_saliva<-regaMicrobiome$SalivaMicrobiome$TaxonomyComposition$Phylum
phylum_stool<-regaMicrobiome$StoolMicrobiome$TaxonomyComposition$Phylum
saliva_meta<-subset(paired_meta,Site=="Saliva"&Response!="NE")
stool_meta<-subset(paired_meta,Site=="Stool"&Response!="NE")
phylum_saliva<-phylum_saliva[,which(colnames(phylum_saliva)%in%saliva_meta$Samples)]
phylum_stool<-phylum_stool[,which(colnames(phylum_stool)%in%stool_meta$Samples)]
phylum_saliva<-data.frame(Samples=colnames(phylum_saliva),t(phylum_saliva))
phylum_stool<-data.frame(Samples=colnames(phylum_stool),t(phylum_stool))
phylum_saliva<-merge(select(paired_meta,c(patientID,Response,Samples)),phylum_saliva,by="Samples")
phylum_stool<-merge(select(paired_meta,c(patientID,Response,Samples)),phylum_stool,by="Samples")
phylum_saliva<-melt(phylum_saliva[,-1],id.vars = c("patientID","Response"),variable.name = "Phylum",value.name = "Abundance")
phylum_stool<-melt(phylum_stool[,-1],id.vars = c("patientID","Response"),variable.name = "Phylum",value.name = "Abundance")
phylum_saliva$Site=rep("Saliva",171)
phylum_stool$Site=rep("Stool",171)
df_p<-bind_rows(phylum_saliva,phylum_stool)
test<-subset(df_p,Phylum=="Firmicutes"&Site=="Stool")
g<-test[order(test$Abundance,decreasing = TRUE),]
df_p$patientID<-factor(df_p$patientID,levels=g$patientID)
levels(factor(df_p$Phylum))
df_p$Phylum<-factor(df_p$Phylum,levels = c("Firmicutes","Bacteroidetes","Proteobacteria",
                                           "Fusobacteriota","Actinobacteriota","Campilobacterota",
                                           "Patescibacteria","Spirochaetota","Desulfobacterota",
                                           "Synergistota","Verrucomicrobiota","Others"))

p1<-ggplot(subset(df_p,Site=="Saliva"),aes(patientID,Abundance,fill=Phylum))+
  geom_bar(stat = "identity", width=1)+
  facet_grid(~Response,scales = "free",space = "free")+
  scale_fill_manual(values = col11)+
  theme_few(base_size = 8)+
  theme(axis.text.x = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title.x = element_blank(),
        legend.margin = margin(0.1,unit="pt"),
        legend.key.size=unit(.1,"inches"),
        legend.text.align=0,
        legend.title=element_text(colour="black",size=8,face = "bold"),
        legend.direction = "horizontal",
        legend.box.just="top",
        legend.spacing = unit(0.1,"cm"),
        legend.spacing.y = unit(0.1,"cm"),
        legend.spacing.x =unit(0.1,"cm"),
        legend.box.spacing = unit(0.1,"cm"),
        legend.justification=c(.4,.4),
        legend.position="top")+
  ggtitle("Patients with paired baseline saliva and stool samples")

p2<-ggplot(subset(df_p,Site=="Stool"),aes(patientID,Abundance,fill=Phylum))+
  geom_bar(stat = "identity", width=1)+
  facet_grid(~Response,scales = "free",space = "free")+
  scale_fill_manual(values = col11,guide=FALSE)+
  theme_few(base_size = 8)+
  theme(axis.text.x = element_text(angle = 90,vjust = 1,hjust = 1,size=6),
        strip.text = element_blank(),
        axis.title.x = element_blank())
```        
      
```{r fig.height=6,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="60%",results='hold'}
plot_grid(p1,p2, labels = c("A","B"), ncol =1, nrow = 2)
```


## beta diversity

### meta

```{r warning=F,results= 'hold'}
meta_file<-subset(meta,Response!="NE"&Cycle=="BL")
meta_file$Group<-paste0(meta_file$Site,"_",meta_file$Response)
meta_file$newName<-paste0(meta_file$patientID,"_",meta_file$Cycle)

meta_file_stool<-subset(meta_file,Site=="Stool")
meta_file_saliva<-subset(meta_file,Site=="Saliva")
meta_file_BL<-meta_file
tree_file <- read.tree("../Data/Data/top500_OTU_phylo_tree.tre")
rooted_tree <- midpoint(tree_file)
otu<-fread("../Data/Data/top500_OTU_table_phylo_tree.csv",data.table = F)
otu<-data.frame(row.names = otu$ID,otu[,-1])

otu_stool<-otu[,which(colnames(otu)%in%meta_file_stool$Samples)]
otu_saliva<-otu[,which(colnames(otu)%in%meta_file_saliva$Samples)]
otu_BL<-otu[,which(colnames(otu)%in%meta_file_BL$Samples)]

otu_stool<-data.frame(t(otu_stool[,order(names(otu_stool))]))
otu_saliva<-data.frame(t(otu_saliva[,order(names(otu_saliva))]))
otu_BL<-data.frame(t(otu_BL[,order(names(otu_BL))]))

unifracs_stool<-GUniFrac(otu_stool, rooted_tree, alpha = c(0.0,0.5,1.0))$unifracs
unifracs_saliva<-GUniFrac(otu_saliva, rooted_tree, alpha = c(0.0,0.5,1.0))$unifracs
unifracs_BL<-GUniFrac(otu_BL, rooted_tree, alpha = c(0.0,0.5,1.0))$unifracs

unifract_dist_stool <- unifracs_stool[, , "d_0.5"]# GUniFrac with alpha 0.5
unifract_dist_saliva <- unifracs_saliva[, , "d_0.5"]
unifract_dist_BL <- unifracs_BL[, , "d_0.5"]

stool_fit<-hclust(as.dist(unifract_dist_stool),method = "ward.D2")
saliva_fit<-hclust(as.dist(unifract_dist_saliva),method = "ward.D2")
BL_fit<-hclust(as.dist(unifract_dist_BL),method = "ward.D2")

stool_tree <- as.phylo(stool_fit)
saliva_tree <- as.phylo(saliva_fit)
BL_tree <- as.phylo(BL_fit)

stool_groups<-as.factor(c(meta_file_stool$Group))
saliva_groups<-as.factor(c(meta_file_saliva$Group))
BL_groups<-as.factor(c(meta_file_BL$Group))

plot_color_BL<-rainbow(length(levels(BL_groups)))[BL_groups]
plot_color_stool<-rainbow(length(levels(stool_groups)))[stool_groups]
plot_color_saliva<-rainbow(length(levels(saliva_groups)))[saliva_groups]
```

### anosim: Analysis of Similarities

**Description**
Analysis of similarities (ANOSIM) provides a way to test statistically whether there is a significant difference between two or more groups of sampling units.

```{r fig.height=7,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="60%",results='hold'}
par(mfrow=c(3,1))
plot(anosim(vegdist(otu_stool,method = "bray"), meta_file_stool$Group),xlab="",ylab="",col=c("gray","#8000FFFF","#80FF00FF"))
plot(anosim(vegdist(otu_saliva,method = "bray"), meta_file_saliva$Group),xlab="",ylab="",col=c("gray","#00FFFFFF","#FF0000FF"))
plot(anosim(vegdist(otu_BL,method = "bray"), meta_file_BL$Group),xlab="",ylab="",col=c("gray","#00FFFFFF","#FF0000FF","#8000FFFF","#80FF00FF"))
```

### adonis: Permutational Multivariate Analysis of Variance

** Description**
Analysis of variance using distance matrices — for partitioning distance matrices among sources of variation and fitting linear models (e.g., factors, polynomial regression) to distance matrices; uses a permutation test with pseudo-F ratios.

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
adonis_stool<-adonis2(as.dist(unifract_dist_stool) ~ stool_groups)
adonis_saliva<-adonis2(as.dist(unifract_dist_saliva) ~ saliva_groups)
adonis_BL<-adonis2(as.dist(unifract_dist_BL) ~ BL_groups)

# Calculate and display the NMDS plot (Non-metric Multidimensional Scaling plot)
res_stool <- metaMDS(unifract_dist_stool,k = 2,distance = "bray")
s.class(
  res_stool$points, col = unique(plot_color_stool), cpoint = 2, 
  fac = stool_groups,
  sub = paste("NMDS plot of OTU \n (p ",adonis_stool$`Pr(>F)`[1],", stress ",
              round(res_stool$stress,3),")",sep="")
)

res_saliva <- metaMDS(unifract_dist_saliva,k = 2,distance = "bray")
s.class(
  res_saliva$points, col = unique(plot_color_saliva), cpoint = 2, 
  fac = saliva_groups,
  sub = paste("NMDS plot of OTU \n (p ",adonis_saliva$`Pr(>F)`[1],", stress ",round(res_saliva$stress,3),")",sep="")
)

res_BL <- metaMDS(unifract_dist_BL,k = 2,distance = "bray")
s.class(
  res_BL$points, col = unique(plot_color_BL), cpoint = 2, 
  fac = BL_groups,
  sub = paste("NMDS plot of OTU \n (p ",adonis_BL$`Pr(>F)`[1],", stress ",
              round(res_BL$stress,3),")",sep="")
)
```
