---
sort: 5
title: "AutoDecoder"
author: "XiaZhao"
date: "4/17/2021"
output: 
  md_document: 
    df_print: kable
    number_sections: yes
    toc: yes
---

[`Return`](./)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Requires

<details>
<summary>
<font size=4>Requires</font>
</summary>
```{r tidy=TRUE,results='hold',message=FALSE}
library(data.table)
library(reshape2)
library(ggthemes)
library(ggsci)
library(tidyverse)
library(FactoMineR)
library(corrplot)
library(colortools)
library(visibly)
library(plotly)
library(scico)
library(factoextra)
library(randomForest)
library(ANN2)
library(NeuralNetTools)
library(ConsensusClusterPlus)
library(survminer)
library(survival)
library(ggExtra)
library(cowplot)
library(corrplot)
library(limma)
source("../R_function/colors.R")
theme_set(theme_cowplot())
"%ni%"<-Negate("%in%")
options(stringsAsFactors = F)
```
</details>


# 8 patients with BL and treat stool and saliva

``````{r warning=F,results= 'hold'}
paired_stool_saliva_TB<-fread("../Data/Data/paired_stool_saliva_TB_8patients.csv",data.table = F)
clinical<-fread("../Data/Data/clinical_adver_41p.csv",data.table = F)[,-10]
otu<-fread("../Data/Data/OTUtable_ori.csv",data.table = F)
colnames(otu)[1]="OTUid"
group<-paired_stool_saliva_TB[,c(1,2)]
group<-group[-which(duplicated(group$patientID)),]
group<-data.frame(row.names = group$patientID,Response=group$Response)
paired_stool_saliva_TB<-paired_stool_saliva_TB[,-2]
paired_stool_saliva_TB$Group<-paste0(paired_stool_saliva_TB$Group,"_",paired_stool_saliva_TB$Site)
paired_stool_saliva_TB<-paired_stool_saliva_TB[,-4]
OTU_sliva_stool_TB<-otu[,which(colnames(otu)%in%c("OTUid",paired_stool_saliva_TB$Samples))]
OTU_sliva_stool_TB<-data.frame(row.names = OTU_sliva_stool_TB$OTUid,OTU_sliva_stool_TB[,-1])
OTU_sliva_stool_TB<-data.frame(Samples=colnames(OTU_sliva_stool_TB),t(OTU_sliva_stool_TB))
OTU_sliva_stool_TB<-merge(paired_stool_saliva_TB,OTU_sliva_stool_TB,by="Samples")
levels(factor(OTU_sliva_stool_TB$Group))
OTU_BL_Saliva<-subset(OTU_sliva_stool_TB,Group=="BL_Saliva")[,-c(1,3)]
OTU_BL_Stool<-subset(OTU_sliva_stool_TB,Group=="BL_Stool")[,-c(1,3)]
OTU_Treat_Saliva<-subset(OTU_sliva_stool_TB,Group=="Treat_Saliva")[,-c(1,3)]
OTU_Treat_Stool<-subset(OTU_sliva_stool_TB,Group=="Treat_Stool")[,-c(1,3)]

OTU_BL_Saliva_stat<-data.frame(OTUid=colnames(OTU_BL_Saliva)[-1],AV=apply(OTU_BL_Saliva[,-1], 2, mean))%>%
  filter(.,AV>5)
OTU_BL_Stool_stat<-data.frame(OTUid=colnames(OTU_BL_Stool)[-1],AV=apply(OTU_BL_Stool[,-1], 2, mean))%>%
  filter(.,AV>5)
OTU_Treat_Saliva_stat<-data.frame(OTUid=colnames(OTU_Treat_Saliva)[-1],AV=apply(OTU_Treat_Saliva[,-1], 2, mean))%>%
  filter(.,AV>5)
OTU_Treat_Stool_stat<-data.frame(OTUid=colnames(OTU_Treat_Stool)[-1],AV=apply(OTU_Treat_Stool[,-1], 2, mean))%>%
  filter(.,AV>5)
OTU_BL_Saliva<-OTU_BL_Saliva[,which(colnames(OTU_BL_Saliva)%in%c("patientID",levels(factor(OTU_BL_Saliva_stat$OTUid))))]
OTU_BL_Stool<-OTU_BL_Stool[,which(colnames(OTU_BL_Stool)%in%c("patientID",levels(factor(OTU_BL_Stool_stat$OTUid))))]
OTU_Treat_Saliva<-OTU_Treat_Saliva[,which(colnames(OTU_Treat_Saliva)%in%c("patientID",levels(factor(OTU_Treat_Saliva_stat$OTUid))))]
OTU_Treat_Stool<-OTU_Treat_Stool[,which(colnames(OTU_Treat_Stool)%in%c("patientID",levels(factor(OTU_Treat_Stool_stat$OTUid))))]
colnames(OTU_BL_Saliva)[-1]<-paste0(colnames(OTU_BL_Saliva)[-1],"_BL_Saliva")
colnames(OTU_BL_Stool)[-1]<-paste0(colnames(OTU_BL_Stool)[-1],"_BL_Stool")
colnames(OTU_Treat_Saliva)[-1]<-paste0(colnames(OTU_Treat_Saliva)[-1],"_Treat_Saliva")
colnames(OTU_Treat_Stool)[-1]<-paste0(colnames(OTU_Treat_Stool)[-1],"_Treat_Stool")
OTU_BL_Saliva_mat<-data.frame(row.names =OTU_BL_Saliva$patientID,OTU_BL_Saliva[,-1])%>%as.matrix()
OTU_BL_Stool_mat<-data.frame(row.names =OTU_BL_Stool$patientID,OTU_BL_Stool[,-1])%>%as.matrix()
OTU_Treat_Saliva_mat<-data.frame(row.names =OTU_Treat_Saliva$patientID,OTU_Treat_Saliva[,-1])%>%as.matrix()
OTU_Treat_Stool_mat<-data.frame(row.names =OTU_Treat_Stool$patientID,OTU_Treat_Stool[,-1])%>%as.matrix()
clinical_mat<-subset(clinical,patientID%in%paired_stool_saliva_TB$patientID)
clinical_mat<-data.frame(row.names =clinical_mat$patientID,clinical_mat[,-1])%>%as.matrix()

OTU8paired_mat<-do.call(cbind,list(clinical_mat,OTU_BL_Saliva_mat,OTU_BL_Stool_mat,
                                   OTU_Treat_Saliva_mat,OTU_Treat_Stool_mat))

res.hc1 <- eclust(OTU_BL_Saliva_mat, "hclust", k = 2,
                  method = "ward.D2", graph =T) 
res.hc2 <- eclust(OTU_BL_Stool_mat, "hclust", k = 2,
                  method = "ward.D2", graph =T) 
res.hc3 <- eclust(OTU_Treat_Saliva_mat, "hclust", k = 2,
                  method = "ward.D2", graph =T) 
res.hc4 <- eclust(OTU_Treat_Stool_mat, "hclust", k = 2,
                  method = "ward.D2", graph =T) 
res.hc5 <- eclust(clinical_mat, "hclust", k = 2,
                  method = "ward.D2", graph =T) 

p1<-fviz_dend(res.hc1, k = 2,main = "BL_Saliva",ggtheme = theme_few(base_size = 6),palette = "lancet",cex = 0.5)

p2<-fviz_dend(res.hc2, k = 2, main = "BL_Stool",ggtheme = theme_few(base_size = 6),palette = "lancet", cex = 0.5)

p3<-fviz_dend(res.hc3, k = 2,main = "Treat_Saliva",ggtheme = theme_few(base_size = 6),palette = "lancet", cex = 0.5)

p4<-fviz_dend(res.hc4, k = 2, main = "BL_Treat",ggtheme = theme_few(base_size = 6),palette ="lancet", cex = 0.5)

p5<-fviz_dend(res.hc5, k = 2,main = "Clinical", ggtheme = theme_few(base_size = 6),palette = "lancet", cex = 0.5)
```

```{r fig.height=5,fig.width=8,dpi=600,fig.align='center',warning=F,out.width ="80%",results='hold'}
plot_grid(p1,p2,p3,p4,p5, labels = c("A","B","C","D","E"), ncol =5 ,nrow = 1)
```

```{r warning=F,results= 'hold'}
load("../Data/Data/regaMicrcobiome.RData")
patients8<-fread("../Data/Data/paired_stool_saliva_TB_8patients.csv",data.table = F)
phylum_stool<-regaMicrobiome$StoolMicrobiome$TaxonomyComposition$Phylum
phylum_saliva<-regaMicrobiome$SalivaMicrobiome$TaxonomyComposition$Phylum
family_stool<-regaMicrobiome$StoolMicrobiome$TaxonomyComposition$Family
family_saliva<-regaMicrobiome$SalivaMicrobiome$TaxonomyComposition$Family

data_list<-list(phylum_stool=phylum_stool,phylum_saliva=phylum_saliva,
                family_stool=family_stool,family_saliva=family_saliva)
data_list<-lapply(data_list, function(x){
  x<-x[,which(colnames(x)%in%patients8$Samples)]
  x<-data.frame(Samples=colnames(x),t(x))
  x<-merge(patients8,x,by="Samples")[,-1]
  x<-melt(x,id.vars = c("patientID" ,"Response","Group","Site" ),
          variable.name = "MicroName",
          value.name = "Abundance")
})

data_list[[1]]$patientID<-factor(data_list[[1]]$patientID,
                                 levels = c("Patient38","Patient41","Patient10",
                                            "Patient13","Patient16","Patient26","Patient32",
                                            "Patient35"))
data_list[[2]]$patientID<-factor(data_list[[2]]$patientID,
                                 levels = c("Patient38","Patient41","Patient10",
                                            "Patient13","Patient16","Patient26","Patient32",
                                            "Patient35"))
data_list[[3]]$patientID<-factor(data_list[[3]]$patientID,
                                 levels = c("Patient38","Patient41","Patient10",
                                            "Patient13","Patient16","Patient26","Patient32",
                                            "Patient35"))
data_list[[4]]$patientID<-factor(data_list[[4]]$patientID,
                                 levels = c("Patient38","Patient41","Patient10",
                                            "Patient13","Patient16","Patient26","Patient32",
                                            "Patient35"))
p1<-ggplot(data_list$phylum_saliva,aes(Group,Abundance,fill=MicroName))+
  geom_bar(stat = "identity", width=1)+
  facet_grid(Site~patientID,space="free",scales = "free")+
  theme_few(base_size = 6)+
  scale_fill_manual(name="Phylum",values = col16)+
  theme(legend.box.just="top",
        legend.spacing = unit(0.1,"cm"),
        legend.spacing.y = unit(0.1,"cm"),
        legend.spacing.x =unit(0.1,"cm"),
        legend.box.spacing = unit(0.1,"cm"),
        legend.justification=c(.4,.4),
        legend.position="top",legend.key.size=unit(.1,"inches"),axis.text.x = element_text(size=6,angle = 90,vjust = 1,hjust = 1),
        axis.title.x = element_blank())
p2<-ggplot(data_list$phylum_stool,aes(Group,Abundance,fill=MicroName))+
  geom_bar(stat = "identity", width=1)+
  facet_grid(Site~patientID,space="free",scales = "free")+
  theme_few(base_size = 6)+
  scale_fill_manual(name="Phylum",values = col16)+
  theme(legend.box.just="top",
        legend.spacing = unit(0.1,"cm"),
        legend.spacing.y = unit(0.1,"cm"),
        legend.spacing.x =unit(0.1,"cm"),
        legend.box.spacing = unit(0.1,"cm"),
        legend.justification=c(.4,.4),
        legend.position="top",legend.key.size=unit(.1,"inches"),axis.text.x = element_text(size=6,angle = 90,vjust = 1,hjust = 1),
        axis.title.x = element_blank())

p3<-ggplot(data_list$family_saliva,aes(Group,Abundance,fill=MicroName))+
  geom_bar(stat = "identity", width=1)+
  facet_grid(Site~patientID,space="free",scales = "free")+
  theme_few(base_size = 6)+
  scale_fill_manual(name="Family",values = col31[c(1:21,24)])+
  theme(legend.box.just="top",
        legend.spacing = unit(0.1,"cm"),
        legend.spacing.y = unit(0.1,"cm"),
        legend.spacing.x =unit(0.1,"cm"),
        legend.box.spacing = unit(0.1,"cm"),
        legend.justification=c(.4,.4),
        legend.position="top",legend.key.size=unit(.1,"inches"), axis.text.x = element_text(size=6,angle = 90,vjust = 1,hjust = 1),
        axis.title.x = element_blank())
p4<-ggplot(data_list$family_stool,aes(Group,Abundance,fill=MicroName))+
  geom_bar(stat = "identity", width=1)+
  facet_grid(Site~patientID,space="free",scales = "free")+
  theme_few(base_size = 6)+
  scale_fill_manual(name="Family",values =col31[c(1:21,24)])+
  theme(legend.box.just="top",
        legend.spacing = unit(0.1,"cm"),
        legend.spacing.y = unit(0.1,"cm"),
        legend.spacing.x =unit(0.1,"cm"),
        legend.box.spacing = unit(0.1,"cm"),
        legend.justification=c(.4,.4),
        legend.position="top",legend.key.size=unit(.1,"inches"),axis.text.x = element_text(size=6,angle = 90,vjust = 1,hjust = 1),
        axis.title.x = element_blank())
```


```{r fig.height=7,fig.width=7,dpi=600,fig.align='center',warning=F,out.width ="100%",results='hold'}
plot_grid(p1,p3,p2,p4, labels = c("A","B","C","D"), ncol =2 ,nrow = 2)
```

```{r warning=F,results= 'hold'}
clinical<-fread("../Data/Data/clinical_adver_41p.csv",data.table = F)
cd<-fread("../Data/Data/cd3cd8.csv",data.table = F)
clinical_cd<-merge(clinical,cd,by="patientID")
treat_BL<-fread("../Data/Data/16pt_BL_treat_pairs_stool.csv",data.table = F)
treat_BL<-subset(treat_BL,patientID%in%clinical_cd$patientID)
treat_BL_BL<-subset(treat_BL,Group=="BL")
treat_BL_Treat<-subset(treat_BL,Group=="Treat")
clinical_cd<-subset(clinical_cd,patientID%in%treat_BL$patientID)
clinical_cd_mat<-data.frame(row.names = clinical_cd$patientID,clinical_cd[,-1])
otu<-regaMicrobiome$StoolMicrobiome$TaxonomyReads$OTU
colnames(otu)[1]="OTUid"
otu_BL<-otu[,which(colnames(otu)%in%c("OTUid",treat_BL_BL$Samples))]
otu_Treat<-otu[,which(colnames(otu)%in%c("OTUid",treat_BL_Treat$Samples))]
otu_BL_stat<-data.frame(OTUid=otu_BL$OTUid,num=apply(otu_BL[,-1],1,sum))%>%
  filter(.,num>10)
otu_Treat_stat<-data.frame(OTUid=otu_Treat$OTUid,num=apply(otu_Treat[,-1],1,sum))%>%
  filter(.,num>10)
otu_BL<-subset(otu_BL,OTUid%in%otu_BL_stat$OTUid)
otu_Treat<-subset(otu_Treat,OTUid%in%otu_Treat_stat$OTUid)

otu_BL_mat<-data.frame(row.names = otu_BL$OTUid,otu_BL[,-1])
otu_Treat_mat<-data.frame(row.names =otu_Treat$OTUid,otu_Treat[,-1])
otu_BL_mat<-data.frame(Samples=colnames(otu_BL_mat),t(otu_BL_mat))
otu_Treat_mat<-data.frame(Samples=colnames(otu_Treat_mat),t(otu_Treat_mat))
otu_BL_mat<-merge(dplyr::select(treat_BL,c(Samples,patientID)),otu_BL_mat,by="Samples")[,-1]
otu_Treat_mat<-merge(dplyr::select(treat_BL,c(Samples,patientID)),otu_Treat_mat,by="Samples")[,-1]
otu_BL_mat<-data.frame(row.names = otu_BL_mat$patientID,otu_BL_mat[,-1])%>%as.matrix()
otu_Treat_mat<-data.frame(row.names = otu_Treat_mat$patientID,otu_Treat_mat[,-1])%>%as.matrix()

colnames(otu_BL_mat)<-paste0(colnames(otu_BL_mat),"_BL")
colnames(otu_Treat_mat)<-paste0(colnames(otu_Treat_mat),"_Treat")
merged_cli_otu<-do.call(cbind,args = list(clinical_cd_mat,otu_BL_mat,otu_Treat_mat))
merged_otu<-do.call(cbind,args = list(otu_BL_mat,otu_Treat_mat))
cd_13<-data.frame(row.names = cd$patientID,cd[,-1])
cd_13<-cd_13[which(rownames(cd_13)%in%rownames(otu_BL_mat)),]

dfcol<-treat_BL[-which(duplicated(treat_BL$patientID)),]
dfcol<-data.frame(row.names = dfcol$patientID,Group=dfcol$Response)
df<-cbind(dfcol,merged_cli_otu)
df$Group<-ifelse(df$Group=="R",1,2)
X<-df[,-c(1:5)]%>%as.matrix()
X<-data.frame(row.names =rownames(X),apply(X, 2, as.numeric))

AE <- autoencoder(X, c(100,10,100), random.seed=1234,
                  loss.type = 'huber',drop.last = F,
                  activ.functions = c('tanh','linear','tanh'),
                  batch.size =3, optim.type = 'rmsprop',
                  n.epochs = 1000, val.prop = 0)

recX <- reconstruct(AE, X)
sort(recX$anomaly_scores, decreasing = TRUE)[1:5]
AE_df<-recX$reconstructed
rownames(AE_df)<-rownames(X)

```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
fviz_nbclust(
  AE_df, 
  kmeans, 
  k.max = 10,
  method = "wss",
  verbose = FALSE)
```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
km.res <- kmeans(scale(AE_df),6, nstart = 25)
fviz_cluster(km.res, data = AE_df,repel = T,labelsize = 12,
            # palette = c("#00AFBB", "#E7B800", "black","red","blue","gray"),
             ggtheme = theme_minimal(),
             main = "Kmeans Clustering Plot")
```

```{r warning=F,results= 'hold'}
cluster<-data.frame(km.res$cluster)
cluster$patientID<-rownames(cluster)
cluster$km.res.cluster<-paste0("km",cluster$km.res.cluster)
stat_num<-cluster%>%group_by(km.res.cluster)%>%summarise(Num=n())
cluster$km.res.cluster<-ifelse(cluster$km.res.cluster==stat_num$km.res.cluster[which(stat_num$Num>5)],"km1","km2")
new<-merge(cluster,clinical,by="patientID")
```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
fit1<-survfit(Surv(PFStime,PFS) ~ km.res.cluster,
              data = new)
fit1
p1<-ggsurvplot(fit1, data=new,pval.method = T,add.all = T,
               tables.theme = theme_classic2(base_size = 8),
               palette  = c("black", "#00AFBB","#E7B800"),
               risk.table = T,
               pval = TRUE,
               legend.title="K-means",
               risk.table.col = "strata",
               surv.median.line = "hv",
               risk.table.y.text.col = T,
               risk.table.y.text = F )
p1
```

```{r warning=F,results= 'hold'}
otu_info<-fread("../Data/Data/OTUtabale_regaStool.csv")[,c(1:7)]
colnames(otu_info)[7]="OTUid"
otu_info$Taxonomy<-paste(otu_info$Phylum,otu_info$Class,otu_info$Order,otu_info$Family,otu_info$Genus,otu_info$Species,sep = "|")
otu_info<-otu_info[,7:8]
ano_data<-X
ano_data$patientID=rownames(ano_data)
ano_data<-merge(cluster,ano_data,by="patientID")
data<-ano_data
data<-data[,-1]
lev<-unique(data$km.res.cluster)
f <- factor(data$km.res.cluster, levels=lev) 
design <- model.matrix(~0+f)
colnames(design) <- lev
eset<-dplyr::select(data,-km.res.cluster)
eset<-data.frame(t(eset))
#eset<-data.frame(apply(eset, 2, av))
cont.wt <- makeContrasts("km1-km2",
                         levels=design) 
fit <- lmFit(eset, design)
fit2 <- contrasts.fit(fit, cont.wt) 
fit2 <- eBayes(fit2) 
tT=topTable(fit2, adjust="BH",sort.by="logFC",n=Inf)
tT = subset(tT, select=c("adj.P.Val","P.Value","logFC"))
colnames(tT)=c("FDR","P.Value","logFC")
range(tT$logFC)
limma_res<-filter(tT,P.Value<=0.05&abs(logFC)>1)
limma_res$Factor<-rownames(limma_res)
limma_res$OTUid<-rownames(limma_res)
limma_res$OTUid<-gsub("_BL","",limma_res$OTUid)
limma_res$OTUid<-gsub("_Treat","",limma_res$OTUid)
limma_res$Group<-rownames(limma_res)
limma_res$Group<-gsub("^.*_","",limma_res$Group)
limma_res<-merge(limma_res,otu_info,by="OTUid")
limma_res$Taxonomy<-gsub("d__Bacteria;k__norank_d__Bacteria;p__","",limma_res$Taxonomy)

knitr::kable(limma_res)

final_res<-dplyr::select(ano_data,c(patientID,km.res.cluster,limma_res$Factor))
final_mat<-data.frame(row.names = final_res$patientID,final_res[,-c(1,2)])%>%as.matrix()
final_col<-data.frame(row.names =ano_data$patientID,Class=ano_data$km.res.cluster)


```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
limma_pca <- PCA(final_mat, graph = FALSE)
fviz_pca_ind(limma_pca,
             label = "none", # hide individual labels
             #habillage = final_col$Class, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE # Concentration ellipses
)
```

```{r warning=F,results= 'hold'}
cli<-fread("../Data/Data/final_clinical_40pt.csv")
meta<-fread("../Data/Data/meta.csv",data.table = F)%>%subset(.,Site=="Stool")
meta<-merge(dplyr::select(meta,c("patientID","Samples","Cycle")),dplyr::select(cli,c("patientID","PFS","PFStime")),by="patientID")
colnames(meta)[2]="Samples"
meta$Group<-ifelse(meta$Cycle=="BL","BL","Treat")
meta_bl<-subset(meta,Group=="BL")
meta_treat<-subset(meta,Group=="Treat")
otu<-regaMicrobiome$StoolMicrobiome$TaxonomyReads$OTU
colnames(otu)[1]="OTUid"
limma_res_BL<-subset(limma_res,Group=="BL")
limma_res_treat<-subset(limma_res,Group=="Treat")
otu<-subset(otu,OTUid%in%limma_res$OTUid)

otu_bl<-otu[,which(colnames(otu)%in%c("OTUid",meta_bl$Samples))]
otu_treat<-otu[,which(colnames(otu)%in%c("OTUid",meta_treat$Samples))]
otu_bl_limma<-subset(otu_bl,OTUid%in%limma_res_BL$OTUid)
otu_treat_limma<-subset(otu_treat,OTUid%in%limma_res$OTUid)
otu_bl_limma<-data.frame(row.names = otu_bl_limma$OTUid,otu_bl_limma[,-1])
otu_bl_limma<-data.frame(Samples=colnames(otu_bl_limma),t(otu_bl_limma))

otu_treat_limma<-data.frame(row.names = otu_treat_limma$OTUid,otu_treat_limma[,-1])
otu_treat_limma<-data.frame(Samples=colnames(otu_treat_limma),t(otu_treat_limma))


df_bl<-merge(meta_bl,otu_bl_limma,by="Samples")
df_treat<-merge(meta_treat,otu_treat_limma,by="Samples")

otu_bl_limma_mat<-data.frame(row.names = df_bl$patientID,df_bl[,-c(1:6)])
otu_treat_limma_mat<-data.frame(row.names = paste0(df_treat$patientID,"_",df_treat$Cycle),df_treat[,-c(1:6)])
```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
fviz_nbclust(
  otu_bl_limma_mat, 
  kmeans, 
  k.max = 10,
  method = "wss",
  verbose = FALSE)
```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
km.res <- kmeans(scale(otu_bl_limma_mat),3, nstart = 25)
fviz_cluster(km.res, data = otu_bl_limma_mat,repel = T,labelsize = 12,
             # palette = c("#00AFBB", "#E7B800", "black","red","blue","gray"),
             ggtheme = theme_minimal(),
             main = "Kmeans Clustering Plot")
```

```{r warning=F,results= 'hold'}
cluster_BL<-data.frame(km.res$cluster)
cluster_BL$patientID<-rownames(cluster_BL)
cluster_BL$km.res.cluster<-paste0("km",cluster_BL$km.res.cluster)
stat_num<-cluster_BL%>%group_by(km.res.cluster)%>%summarise(Num=n())
cluster_BL$km.res.cluster<-ifelse(cluster_BL$km.res.cluster==stat_num$km.res.cluster[which(stat_num$Num>5)],"km1","km2")
```

```{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="30%",results='hold'}
fviz_nbclust(
  otu_treat_limma_mat, 
  kmeans, 
  k.max = 10,
  method = "wss",
  verbose = FALSE)
```

```{r warning=F,results= 'hold'}
set.seed(123)
km.res <- kmeans(scale(otu_treat_limma_mat),4, nstart = 25)
fviz_cluster(km.res, data = otu_treat_limma_mat,repel = T,labelsize = 12,
              palette = c("#00AFBB", "#E7B800", "black","red","blue","gray"),
             ggtheme = theme_minimal(),
             main = "Kmeans Clustering Plot")
cluster_treat<-data.frame(km.res$cluster)
cluster_treat$patientID<-rownames(cluster_treat)
cluster_treat$km.res.cluster<-paste0("km",cluster_treat$km.res.cluster)
stat_num<-cluster_treat%>%group_by(km.res.cluster)%>%summarise(Num=n())
knitr::kable(stat_num)
cluster_treat$km.res.cluster<-gsub("km4","km2",cluster_treat$km.res.cluster)
df_bl<-merge(cluster_BL,df_bl,by="patientID")
df_treat$patientID<-paste0(df_treat$patientID,"_",df_treat$Cycle)
df_treat<-merge(cluster_treat,df_treat,by="patientID")


fit1<-survfit(Surv(PFStime,PFS) ~ km.res.cluster,
              data = df_bl)
fit1
fit2<-survfit(Surv(PFStime,PFS) ~ km.res.cluster,
              data = df_treat)
fit2
```

``````{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="50%",results='hold'}
p1<-ggsurvplot(fit1, data=df_bl,pval.method = T,combine = T, 
               palette  = c("black", "#00AFBB","#E7B800","red","green"),
               tables.theme = theme_bw(base_size = 8),
               risk.table = T,
               pval = TRUE,
               ggtheme = theme_survminer(),
               legend.title="k-means",
               risk.table.col = "strata",
               surv.median.line = "hv",
               risk.table.y.text.col = T,
               risk.table.y.text = F )
p1

```


``````{r fig.height=5,fig.width=5,dpi=600,fig.align='center',warning=F,out.width ="50%",results='hold'}
p2<-ggsurvplot(fit2, data=df_treat,pval.method = T,combine = T, 
               palette  = c("black","#00AFBB","#E7B800","red"),
               tables.theme = theme_bw(base_size = 8),
               risk.table = T,
               pval = TRUE,
               ggtheme = theme_survminer(),
               legend.title="k-means",
               risk.table.col = "strata",
               surv.median.line = "hv",
               risk.table.y.text.col = T,
               risk.table.y.text = F )
p2
```

